### 2012-06-13 trunk/ijcis-pos-model/sources/temporalModel.tex
version sent to Olga and Nicolas

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on the GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated with an example. The information in the system is defined by the following elements:

\begin{definition}
\label{def:generalized-fuzzy-domain}
\textbf{Generalized fuzzy domain.}
$D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
\begin{equation}
\label{eq:generalized-fuzzy-domain}
D_G \subseteq \tilde \Pow\left(D \right)\cup \text{NULL}
\end{equation}
\end{definition}
The datatypes for $D_G$ are shown in table \ref{tbl:gefred-data-types}.

\begin{definition}
\label{def:typeof-domain}
\textbf{Typeof(D).}
The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$. For datatype, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
\end{definition}

@@

It is possible to define a more specific generalized temporal domain, $\T_G$:

\begin{definition}
\label{def:generalized-fuzzy-temporal-domain}
\textbf{Generalized fuzzy temporal domain.}
If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
The Generalized Fuzzy Temporal Domain, $\T_G$ is
\begin{equation}
\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
\end{equation}
\end{definition}

@@

\begin{definition}
\textbf{Generalized fuzzy temporal relation.}
A generalized fuzzy temporal relation $R_{FTG}$ is given by:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

\begin{example}
\label{ex:sample-ft-database}
 Consider a library database as explained in example \ref{ex:library-database}. But now, the time is stored with imprecision. For simplicity, the ill-known time points are represented as triangular fuzzy numbers. An ill-known time point is given by $[dd/mm/yyyy,a,b]$. The values for $a$ and $b$ are integers. Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
\begin{align}
 \label{eq:head-ex}
\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
\nonumber
&  \left(\text{Title}:D_{\text{title}}\right),\\
\nonumber
& \left(\text{Loan}:D_{\text{Loan}}\right) \\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
\end{align}

@@

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample}Sample library database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c |c c}\\
\hline
& & & \multicolumn{2}{c}{PVP} \\
$\Head$ &  \textbf{ID}  & Title  &  Start Loan & End Loan\\
\hline
\multirow{3}{*}{$\Body$ \Big\lbrace} &  3 & Dracula & [15/3/2012, 5, 2] & [30/3/2012, 1, 1] \\
 &4 & Frankenstein & [25/3/2012, 3, 2] & [4/4/2012, 1, 7] \\
&5 & Harry Potter & [18/3/2012, 4, 1] & [2/4/2012, 2, 2]\\
% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
\hline\\
\end{tabular}
}

\begin{definition}
\label{def:value-component}
\textbf{Value component.}
The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
\begin{align}
\label{eq:value-component}
R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
\nonumber
\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace 
\end{align}
\end{definition}

For example, in the case of the record with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Title}:D_{\text{title}}\right),\\
\nonumber
& \left(\text{Loan}:D_{\text{Loan}}\right), \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \left \lbrace 3:, \text{``Dracula``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
\end{align}

\begin{definition}
\label{def:compatibility-component}
\textbf{Compatibility component.}
The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
\begin{align}
\label{eq:compatibility-component}
R^{c}_{FTG} = \left \lbrace \Head^{c},\Body^{c} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{c} = \left \lbrace \left[C_{A_{G1}} \right], \ldots,  \left[C_{A_{Gn}} \right] \right \rbrace \\
\nonumber
\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace 
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:temporal-component}
\textbf{Temporal component.}
The temporal component $R^{t}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the temporal components for both the head and the body of the relation:
\begin{align}
\label{eq:temporal-component}
R^{t}_{FTG} = \left \lbrace \Head^{t},\Body^{t} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{t} = \left \lbrace \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \right \rbrace \\
\nonumber
\Body^ {t} = \left \lbrace  \left[ \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right]  \right \rbrace 
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:generalized-primary-key}
\textbf{Generalized primary key.}
A generalized primary key, $K_G$ is a subset of the head:
\begin{align}
\label{eq:generalized-primary-key}
K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\textbf{Generalized fuzzy temporal key.}
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
\nonumber
 \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\nonumber
\text{Typeof }\left(D_{ID} \right) = \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c | c c}\\
\hline
& & & &\multicolumn{2}{c}{PVP} \\
$\Head$  & \textbf{ID}  & \textbf{V} & Title  & Start Loan & End Loan \\
\hline
\multirow{4}{*}{$\Body  \Big\lbrace$}  & 3 & 001 & Dracula & [15/3/2012, 5, 2] & [30/3/2012, 1, 1] \\
& 4 & 001 & Frankenstein & [25/3/2012, 3, 2] & [4/4/2012, 1, 7]\\
& 5 & 001 & Harry Potter & [18/3/2012, 4, 1] & [2/4/2012, 2, 2]\\
& 3 & 002 & Dracula & [4/4/2012, 3, 3] & UC \\
\hline\\
\end{tabular}
}

@@

The algebra for the data manipulation is defined in \cite{Medina1994}. The Generalized Fuzzy Relational Algebra manipulates relations like $R_{FG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify, Delete, } and \emph{Revise}. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.

It is important to notice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.  Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$). Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):

@@

In the implementation of the DML operations we will consider a strictly consistent temporal database. This is a natural extension of the crisp temporal databases. This approach allows a easy implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.

@@

\begin{eqnarray}
\label{eq:pvp-close-current}
\text{Close-current} \left(R, \left(A, J \right) \right) =\\
\begin{cases}
\nonumber
\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
\nonumber
\mbox{ if } I = \mbox{ Current } \left(R, A \right) \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber R , \text{ in any other case}
\end{cases}
\end{eqnarray}
\end{definition}

@@

\begin{verbatim}
Insert(3,'Dracula',
[[15/3/2012, 5, 2],
        [30/3/2012, 1, 1]]);
Insert(4,'Frankenstein',
[[25/3/2012, 3, 2], 
        [4/4/2012, 1, 7]]);
Insert(5,'Harry Potter',
[[18/3/2012, 4, 1], 
        [2/4/2012, 2, 2]]);
\end{verbatim}

\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
	\begin{enumerate}
	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$. Note that here, the result of the Overlaps operator is in the unit interval.
For example, the librarian loans the book with ID=3 but the book is not returned yet. The insert sentence is the following.
 \begin{verbatim}
Insert(3,'Dracula',
[[4/4/2012, 3, 3], UC]);
	      \end{verbatim}

	\item Modify and close the current version of the entity $A$ and insert $\left(A, I \right)$. For example, consider now that the book with ID 3 is loaned  around the 24/4/2012 (this is model by a triangular fuzzy number: $\left[24/4/2012, 1, 1 \right]$). Here the problem is that the book with ID 3 was loaned around 4/4/2012, but not returned. If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date. %This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
	    \begin{verbatim}
Insert(3,'Dracula',
[[24/4/2012, 1, 1], UC]]);
	      \end{verbatim}

	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R$. For example, consider now that the librarian wants to introduce a past loan for the book with ID 3. The book was loaned around 6/4/12/2012 and returned around 25/4/2012. As this interval does overlaps other time intervals with a degree of 1, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected. The insert sentence is:

	      \begin{verbatim}
Insert(3,'Dracula',
[[6/4/12/2012, 1, 1],
         [25/4/2012, 1, 1]]);
	    \end{verbatim}
	\end{enumerate}

@@

\begin{align}
\label{eq:insert-fuzzy-temporal}
\INS \left(R ,\left(A, I \right) \right) &=\\
\begin{cases}
\nonumber
R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \vee \\
&  \left(A, J \right) \in R \wedge I \text{Overlaps} J < 1\\
%\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
%\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
%\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
%\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
%&  J<I<K  \big \rbrace\\
\MOD \left(R, \left(A, I \right)\right), & \mbox{ otherwise }  
\end{cases} 	
\end{align}
\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
The delete operation logically removes an entity $A$ which is valid in the relation $R$:
\begin{align}
\label{eq:delete-fuzzy-temporal}
\DEL \left(R ,A \right) =\\
\begin{cases}
\nonumber
R - \left(A, I \right), &  \forall \left( A, I \right) \in R\\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

For example, consider that the librarian wants to delete the history for the book with ID = 3. The following sentence deletes all the versions for the book with ID = 3.

\begin{verbatim}
 Delete(3,'Dracula');
\end{verbatim}

\subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$. This operation is used when a correction on the values for the entity $A$ should be done.
\begin{align}
\label{eq:revise-fuzzy-temporal}
\REV\left(R, \left(A, I \right)\right)=\\
\begin{cases}
\nonumber
 R - \left(A', I \right) \cup \left \lbrace\left(A, I \right)\right \rbrace  & \mbox{ if } \left(A', I \right) \in R \\%: A = \left(a_1, \ldots, a_n \right),\\
%A'= \left(a_1', \ldots, a_n' \right)   \\   
R, & \mbox{ otherwise }
\end{cases}
\end{align}

For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5. The new title is `Harry Potter IV'. The revise sentence is:

\begin{verbatim}
Revise(5,'Harry Potter IV', 
[[18/3/2012, 4, 1], 
          [2/4/2012, 2, 2]]);
\end{verbatim}

>>>>

In this section we will formalize the model for valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, the data manipulation language is defined.

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated with examples. The information in the system is defined by the following elements:

\begin{definition}
\label{def:generalized-fuzzy-domain}
\emph{Generalized fuzzy domain.}
Let $D$ be the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
\begin{equation}
\label{eq:generalized-fuzzy-domain}
D_G \subseteq \tilde \Pow\left(D \right)\cup \text{NULL}
\end{equation}
\end{definition}
The datatypes that can be used to represent $D_G$ are shown in table \ref{tbl:gefred-data-types}.

\begin{definition}
\label{def:typeof-domain}
\textbf{Typeof(A).}
Consider $D_G$ to be a generalized fuzzy domain and let $A \in D_G$ be a value for the domain. 
The function typeof$\left(A \right)$ returns the datatype associated with the domain $D_G$. For a given value $A$, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
\end{definition}

@@

It is possible to define a more specific generalized temporal domain, $\T_G$.

\begin{definition}
\label{def:generalized-fuzzy-temporal-domain}
\emph{Generalized fuzzy temporal domain.}
Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
The Generalized Fuzzy Temporal Domain, $\T_G$ is
\begin{equation}
\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
\end{equation}
\end{definition}

@@

\begin{definition}
\emph{Generalized fuzzy temporal relation.}
A generalized fuzzy temporal relation $R_{FTG}$ is given by~\cite{Medina1994}:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

\begin{example}
\label{ex:sample-ft-database}
 Consider a hospital database as explained in example \ref{ex:library-database}. But now, the time is stored with imprecision, because the date of arrival and leave are not precisely known. For simplicity, the ill-known time points are represented as triangular fuzzy numbers. An ill-known time point is given by $[dd/mm/yyyy,a,b]$. The values for $a$ and $b$ are integers. Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
\begin{align}
 \label{eq:head-ex}
\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
\nonumber
&  \left(\text{Name}:D_{\text{Name}}\right),\\
% \nonumber
% & \left(\text{Loan}:D_{\text{Loan}}\right) \\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
\end{align}

@@

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c |p{0.35\columnwidth}}\\
\hline
& & & \centerline{PVP} \\
$\Head$ &  \textbf{ID}  & Name  &  Arrival , Leave\\
\hline
\multirow{3}{*}{$\Body$ } &  3 & John & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
 &4 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
&5 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
\hline\\
\end{tabular}
}

\begin{definition}
\label{def:value-component}
\emph{Value component.}
The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
\begin{align}
\label{eq:value-component}
R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
\nonumber
\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace 
\end{align}
\end{definition}

For example, in the case of the record with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Title}:D_{\text{title}}\right),\\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \left \lbrace 3:, \text{``John``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
\end{align}

\begin{definition}
\label{def:compatibility-component}
\emph{Compatibility component.}
The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
\begin{align}
\label{eq:compatibility-component}
R^{c}_{FTG} = \left \lbrace \Head^{c},\Body^{c} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{c} = \left \lbrace \left[C_{A_{G1}} \right], \ldots,  \left[C_{A_{Gn}} \right] \right \rbrace \\
\nonumber
\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace 
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:temporal-component}
\emph{Temporal component.}
The temporal component $R^{t}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the temporal components for both the head and the body of the relation:
\begin{align}
\label{eq:temporal-component}
R^{t}_{FTG} = \left \lbrace \Head^{t},\Body^{t} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{t} = \left \lbrace \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \right \rbrace \\
\nonumber
\Body^ {t} = \left \lbrace  \left[ \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right]  \right \rbrace 
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:generalized-primary-key}
\emph{Generalized primary key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
A generalized primary key, $K_G$ is a subset of the head:
\begin{align}
\label{eq:generalized-primary-key}
K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\emph{Generalized fuzzy temporal key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
\nonumber
 \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\nonumber
\text{Typeof }\left(D_{ID} \right) = \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c | p{0.35\columnwidth} }\\
\hline
% & & & \multicolumn{2}{c}{PVP} \\
%& & & & \centerline{PVP}\\
{$\Head$}  &  &  & &(Arrival\\
& \textbf{ID}  & \textbf{V} & Name& ,Leave )	\\
 \hline
\multirow{4}{*}{$\Body$}  & 3 & 001 & John & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
& 4 & 001 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
& 5 & 001 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
 & 3 & 002 & John & [4/4/2012, 3, 3] , UC \\
\hline\\
\end{tabular}
}

@@

The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify, Delete, }. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
%The algebra for the data manipulation is defined in .

It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.  Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$). Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):

@@

In the implementation of the DML operations we will consider a strictly consistent temporal database. This is a natural extension of the crisp temporal databases. This approach allows an implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.

@@

\begin{eqnarray}
\label{eq:pvp-close-current}
\text{Close-current} \left(R, \left(A, J \right) \right) =\\
\begin{cases}
\nonumber
\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
\nonumber
\mbox{ if Current } \left(R, A \right) \neq \emptyset \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber R , \text{ in any other case}
\end{cases}
\end{eqnarray}
\end{definition}

@@

\begin{verbatim}
Insert(3,'John',
[[15/3/2012, 5, 2],
        [30/3/2012, 1, 1]]);
Insert(4,'Antoon',
[[25/3/2012, 3, 2], 
        [4/4/2012, 1, 7]]);
Insert(5,'Daan',
[[18/3/2012, 4, 1], 
        [2/4/2012, 2, 2]]);
\end{verbatim}

\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
	\begin{enumerate}
	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$. Note that here, the result of the Overlaps operator is in the unit interval.
For example, the patient with ID=3 has not left the hospital yet. The insert sentence is the following.

 \begin{verbatim}
Insert(3,'John',
[[4/4/2012, 3, 3], UC]);
	      \end{verbatim}

	\item Modify and close the current version of the entity $A$ and insert $\left(A, I \right)$. For example, consider now that the hospital with ID=3 arrived  around the 24/4/2012 (this is model by a triangular fuzzy number: $\left[24/4/2012, 1, 1 \right]$). Here the problem is that the patient with ID=3 arrived around 4/4/2012, but it was not stored in the database the leave date. If the patient is again in the hospital, then it is necessary to set the leave date and add a new row with the new arrival date. %This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
	    \begin{verbatim}
Insert(3,'John',
[[24/4/2012, 1, 1], UC]]);
	      \end{verbatim}

	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R$. For example, consider now that the hospital manager wants to introduce a past arrival for the patient with ID=3. The patient arrived around 6/4/12/2012 and left the hospital around 25/4/2012. As this interval does overlaps other time intervals with a degree of 1, it is not possible that the patient was it the hospital at that time interval. Therefore, the insertion is rejected. The insert sentence is:

	      \begin{verbatim}
Insert(3,'John',
[[6/4/12/2012, 1, 1],
         [25/4/2012, 1, 1]]);
	    \end{verbatim}
	\end{enumerate}

@@

\begin{align}
\label{eq:insert-fuzzy-temporal}
\INS \left(R ,\left(A, I \right) \right) &=\\
\begin{cases}
\nonumber
R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \mbox{ or }\\
&  \left(A, J \right) \in R \mbox{ and } I \text{ Overlaps } J < 1\\
%\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
%\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
%\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
%\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
%&  J<I<K  \big \rbrace\\
\MOD \left(R, \left(A, I \right)\right), & \mbox{ otherwise }  
\end{cases} 	
\end{align}
\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
The delete operation logically removes an entity $A$ which is valid in the relation $R$:
\begin{align}
\label{eq:delete-fuzzy-temporal}
\DEL \left(R ,A \right) =\\
\begin{cases}
\nonumber
R - \left \lbrace \left(A, I_1 \right), \ldots, \left(A, I_n \right) \right \rbrace, &  \forall I_i: \left( A, I_i \right) \in R\\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The following sentence deletes all the rows for the patient with ID = 3.

\begin{verbatim}
 Delete(3,'John');
\end{verbatim}

% \subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
% This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$. This operation is used when a correction on the values for the entity $A$ should be done.
% \begin{align}
% \label{eq:revise-fuzzy-temporal}
% \REV\left(R, \left(A, I \right)\right)=\\
% \begin{cases}
% \nonumber
%  R - \left(A', I \right) \cup \left \lbrace\left(A, I \right)\right \rbrace  & \mbox{ if } \left(A', I \right) \in R \\%: A = \left(a_1, \ldots, a_n \right),\\
% %A'= \left(a_1', \ldots, a_n' \right)   \\   
% R, & \mbox{ otherwise }
% \end{cases}
% \end{align} 
% 
% 
% For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5. The new title is `Harry Potter IV'. The revise sentence is:
% 
% \begin{verbatim}
% Revise(5,'Harry Potter IV', 
% [[18/3/2012, 4, 1], 
%           [2/4/2012, 2, 2]]);
% \end{verbatim}
